---
title: "Extra Galactic Background Light"
author: "Aaron Robotham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extra Galactic Background Light}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The Extra Galactic Background Light (EBL) is the flux density observed across all wavelengths of light. It is dominated by the CMB, followed by the optical and NIR (about 10-20\%). There are a few ways to measure it, but my work has focussed on the method of integrating out the number counts of galaxies at different wavelengths. Assuming we have sufficiently deep photometry, this should directly capture most of the EBL, with the rest estimated by either a model or non-parametric spline fit (see Driver et al 2016).

## Generating the EBL

You will need ProSpect v0.8.6 for this vignette to fully run.

Load the libraries we need:

```{r}
library(ProSpect)
library(celestial)
library(foreach)
library(magicaxis)
#library(doParallel)
#registerDoParallel(cores=detectCores()-2)
```

Encode the Madua and Dickenson 2014 CSFH function:

```{r}
MD14func = function(z, norm=0.00945){ #norm here is MD14 0.015*0.63 (latter Salp to Chab Driver 2016 correction)
  norm * ((1+z)^2.7) / ((1+((1+z)/2.9)^5.6))
}
```

Create a function to generate the EBL:

```{r}
EBL = function(massfunc_z = MD14func, norm=0.00945, Zstart=1e-4, Zfinal=0.02,
               zvec = 10^seq(-3,1, by=0.1), SFRvec=rep(0.01,length(zvec)),
               filters=c('FUV_GALEX', 'NUV_GALEX', 'u_SDSS', 'g_SDSS', 'r_SDSS', 'i_SDSS',
               'Z_VISTA', 'Y_VISTA', 'J_VISTA', 'H_VISTA', 'K_VISTA', 'W1_WISE' , 'W2_WISE',
               'W3_WISE', 'W4_WISE', 'P100_Herschel', 'P160_Herschel', 'S250_Herschel' ,
               'S350_Herschel', 'S500_Herschel'), out='Photom', ...){
  
  data(Dale_NormTot)
  data(BC03lr)
  agevec = cosdistTravelTime(zvec, ref='planck') * 1e9 # Create age vector
  
  if(!is.null(massfunc_z)){ # You can either pass in an SFR function, or an SFRvec of the values at redshifts given by zvec
    SFRvec = massfunc_z(zvec, norm=norm)
  }
  
  covolvec = cosdistCoVol(zvec, OmegaM = 0.3, OmegaL=0.7, H0=70) * 1e9 #  x1e9 to get to Mpc^3 (output is Gpc^3)
  volweights = (c(0,diff(covolvec)) + c(diff(covolvec),0))/2 # This evens out the volume weights around bin edges
  filtout = foreach(i = filters)%do%{approxfun(getfilt(i))} # Generate our filter sets for quick generation
  
  tempSFH = approxfun(agevec, SFRvec, yleft = 0, yright = 0) # Generate our temporary SFH function to create our ZH
  Zvec = Zfunc_massmap_box(agevec, massfunc=tempSFH, Zstart=Zstart, Zfinal=Zfinal) # Create out closed box ZH
  
  fluxz = foreach(i = 1:length(agevec), .combine='rbind')%do%{
    tempSFH = approxfun(agevec - agevec[i], SFRvec, yleft = 0, yright = 0) # Create a SFH from the current time to the end of the Universe
    tempZH = function(x,...){approxfun(agevec - agevec[i], Zvec, yleft = 0, yright = 0)(x)} # Create a ZH from the current time to the end of the Universe
    
    if(out=='Photom'){
      ProSpectSED(SFH = SFHfunc, massfunc=tempSFH, z = zvec[i], agemax = max(agevec) - agevec[i], filtout=filtout, # Run ProSpect SED for our target tempSFH and tempZH
                Dale=Dale_NormTot, speclib=BC03lr, OmegaM = 0.3, OmegaL=0.7, H0=70, Z=tempZH, returnall=FALSE, ...)
    }else if(out=='Spec'){
      temp=ProSpectSED(SFH = SFHfunc, massfunc=tempSFH, z = zvec[i], agemax = max(agevec) - agevec[i], filters=NULL, # Run ProSpect SED for our target tempSFH and tempZH
                Dale=Dale_NormTot, speclib=BC03lr, OmegaM = 0.3, OmegaL=0.7, H0=70, Z=tempZH, returnall=TRUE, ...)$FinalFlux
      temp=approxfun(temp)
      temp(10^seq(2, 7, by=0.01))
    }
  }
  
  fluxz = fluxz * volweights # Weight by volume elements
  flux = colSums(fluxz, na.rm = TRUE) # Sum epochs
  flux = flux * 1e-26 # x1e-26 to get to W/m2/Hz (output is Jansky)
  flux = flux * 1e9 # to get to nW/m2/Hz
  flux = flux / (4*pi) # to get to nW/m2/Hz/sr
  
  if(out=='Photom'){
    tempcen = cenwave[cenwave$filter %in% filters,'cenwave']
    tempfreq =  299792458/(tempcen/1e10) # Convert central wavelength from wavelength to freq
    flux = flux * tempfreq # Scale to nW/m2/sr
    output=data.frame(wave=tempcen, flux=flux)
  }else if(out=='Spec'){
    tempwave = 10^seq(2, 7, by=0.01)
    tempfreq =  299792458/(tempwave/1e10) # Convert central wavelength from wavelength to freq
    flux = flux * tempfreq # Scale to nW/m2/sr
    output=data.frame(tempwave, flux=flux)
  }
  return(invisible(output))
}
```

The wavelength is in Angstroms and the flux is nW/m2/sr (as per Driver et al 2016).

Now we can generate some EBL models (need to up the default zvec resolution to get this looking smooth though):

```{r, fig.width=8, fig.height=6}
magplot(EBL(out='Spec', zvec = 10^seq(-3,1, by=0.01)), xlab='Wave / Ang', ylab='nW/m^2/sr',
        ylim=c(0,13), log='x', type='l', grid=TRUE)
lines(EBL(norm=0.008, Zfinal=0.02, out='Spec', zvec = 10^seq(-3,1, by=0.01)), col='red')
lines(EBL(norm=0.008, Zfinal=0.01, out='Spec', zvec = 10^seq(-3,1, by=0.01)), col='blue')
lines(EBL(norm=0.008, Zstart=0.02, Zfinal=0.02, out='Spec', zvec = 10^seq(-3,1, by=0.01)), col='darkgreen')
lines(EBL(norm=0.008, Zstart=0.01, Zfinal=0.01, out='Spec', zvec = 10^seq(-3,1, by=0.01)), col='brown')
```

## Fitting the EBL

Can we fit the EBL? Let's try with 5\% error, where we just use optical and NIR filters:

```{r}
set.seed(1)
filters=c('u_SDSS', 'g_SDSS', 'r_SDSS', 'i_SDSS', 'Z_VISTA', 'Y_VISTA', 'J_VISTA',
          'H_VISTA', 'K_VISTA')
EBLin = EBL(norm=0.008, Zfinal=0.01, filters=filters)
EBLin$fluxerr = EBLin$flux*0.05
EBLin$flux = rnorm(9, mean=EBLin$flux, sd=EBLin$fluxerr)
```

Take a look at our mock EBL:

```{r, fig.width=8, fig.height=6}
magplot(EBLin[,1:2], ylim=c(0,13), xlab='Wave / Ang', ylab='nW/m^2/sr', grid=TRUE)
magerr(EBLin$wave, EBLin$flux, ylo=EBLin$fluxerr)
```

Here we will fit for the normalisation and the final metallicity in our closed box model (notice we do the fit in log space):

```{r, eval=FALSE}
# Make likelihood function:
EBLlike=function(par, Data){
  -sum(dnorm(x = EBL(norm=10^par[1], Zfinal=10^par[2], filters=filters)$flux,
             mean = Data$flux, sd = Data$fluxerr, log = TRUE))
}
# Fit it (this takes about a minute):
EBLfit = optim(par=c(-2,-1.7), EBLlike, Data=EBLin, method = 'L-BFGS-B',
               lower=c(-3,-3), upper=c(-1,-1), hessian = TRUE)
```

And now we can check the outputs. Here I basically get back the model we put in:

```{r, eval=FALSE}
print(10^EBLfit$par) # model parameters, I see 0.008235409, 0.012355299
print(sqrt(diag(solve(EBLfit$hessian)))) # dex errors, I see 0.01309136, 0.11251226
```

Compare our fitted model to the observations:

```{r, eval=FALSE, fig.width=8, fig.height=6}
magplot(EBLin[,1:2], ylim=c(0,13), xlab='Wave / Ang', ylab='nW/m^2/sr', grid=TRUE)
magerr(EBLin$wave, EBLin$flux, ylo=EBLin$fluxerr)
lines(EBL(norm=10^EBLfit$par[1], Zfinal=10^EBLfit$par[2]), col='red')
```

## SFH Normalisation

Looking at the EBLs generated above, it is clear that even for very different metallicity evolutions, the K band EBL flux is a very good estimator of the SFH normalisation.

We can check this:

```{r}
set.seed(666)
norm_ran = runif(20, 0.007, 0.015)
Zstart_ran = runif(20, 1e-4, 0.04)
Zfinal_ran = runif(20, Zstart_ran, 0.04)
```

Generate the EBLs:

```{r}
Knorm = foreach(i=1:20, .combine='c')%do%{EBL(norm=norm_ran[i], Zstart = Zstart_ran[i],
                                            Zfinal=Zfinal_ran[i], filters='K_VISTA')[1,2]}
```

Find the best fit line:

```{r}
KvEBL = lm(Knorm ~ norm_ran)
print(KvEBL)
```

Plot the K band EBL flux versus the input normalisation (for very different ZH remember):

```{r, fig.width=8, fig.height=8}
magplot(norm_ran, Knorm, xlab='SFH Norm (Msol/Yr/Mpc^3)', ylab='K band EBL flux (nW/m2^2/sr)',
        grid=TRUE)
abline(KvEBL, col='red')
```