---
title: "Fake It Until You Make It"
author: "Aaron Robotham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mock Fitting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Making Mock Data

Load the libraries we will need:

```{r}
library(ProSpect)
library(LaplacesDemon)
library(foreach)
library(celestial)
library(magicaxis)
```

Load some data we want to use:

```{r}
data("BC03lr") #BC03 spectral library
data("Dale_NormTot") #Normalised Dale templates
data("pivwave") # Pivot/effective wavelengths
```

```{r}
set.seed(1)
redshift = 0.1

filters=c('FUV_GALEX', 'NUV_GALEX', 'u_SDSS', 'g_SDSS', 'r_SDSS', 'i_SDSS', 'Z_VISTA',
          'Y_VISTA', 'J_VISTA', 'H_VISTA', 'K_VISTA', 'W1_WISE' , 'W2_WISE', 'W3_WISE',
          'W4_WISE', 'P100_Herschel', 'P160_Herschel', 'S250_Herschel' , 'S350_Herschel',
          'S500_Herschel')

filtout=foreach(i = filters)%do%{approxfun(getfilt(i))}

temppiv=pivwave[pivwave$filter %in% filters,]

agemax = 13.3e9-cosdistTravelTime(z=redshift, H0 = 67.8, OmegaM = 0.308)*1e9
```

Create our model galaxy We will be using a snorm_trunc SFH function and a massmap_lin metallicity with the following parameters:

```{r}
inpar=c(mSFR = 0, #log-space
        mpeak = 0.7, #log-space
        mperiod = 0.3, #log-space
        mskew = 0.3,
        tau_birth = 0, #log-space
        tau_screen = -0.5, #log-space
        alpha_SF_birth = 1,
        alpha_SF_screen = 3
        )
```

Simple function to view our target SFH:

```{r}
plotSFH=function(par,agemax=13.3, add=FALSE,col='black',ylim=NULL,...){
 magcurve(massfunc_snorm_trunc(age=x,mSFR=10^par[1],mpeak=10^par[2],mperiod=10^par[3],
                               mskew=par[4], magemax=agemax),0,13.8e9,add=add,col=col,
                               ylim=ylim,xlab='Age (Yr)', ylab='SFR (Msol / Yr)',...)
}
```

Let's take a look:

```{r, fig.width=8, fig.height=6}
plotSFH(inpar)
```

Now we can generate our galaxies SED:

```{r}
genSED=ProSpectSED(massfunc=massfunc_snorm_trunc,
                   mSFR=10^inpar[1],
                   mpeak=10^inpar[2],
                   mperiod=10^inpar[3],
                   mskew=inpar[4],
                   tau_birth=10^inpar[5], 
                   tau_screen=10^inpar[6], 
                   alpha_SF_birth=inpar[7], 
                   alpha_SF_screen=inpar[8],
                   z=0.1,
                   Z=Zfunc_massmap_lin,
                   filtout=filtout,
                   Dale=Dale_NormTot,
                   speclib=BC03lr,
                   agemax=agemax
)
```

At this point we create our mock photometry catalogue entry with a fractional error in flux of 0.1 assumed (roughly 0.092 mag error). The photom output of **ProSpectSED** is always Jansky, which is what we want for fitting (you shouldn't fit in magnitude space!).

```{r}
flux_input=data.frame(filter=temppiv$filter, pivwave=temppiv$pivwave, flux=genSED$Photom, fluxerr=genSED$Photom*0.1)
print(flux_input)
```

## Fitting a Galaxy

In principle if you have a real galaxy to fit and it is already in the required format of **flux_input**, you can start from this point.

To speed things up we will also pre-compute the Luminosity distance to our source (since this will be the same for every iteration, and will save some computation time):

```{r}
LumDist_Mpc = cosdistLumDist(z=0.1, H0 = 67.8, OmegaM = 0.308)
```

```{r}
Data=list(flux=flux_input,
          arglist=list(z=0.1, massfunc=massfunc_snorm_trunc, agemax=agemax, Z=Zfunc_massmap_lin, LumDist_Mpc=LumDist_Mpc),
          speclib=BC03lr, 
          Dale=Dale_NormTot, 
          filtout=filtout, 
          SFH=SFHfunc, # the preferred functional form of the SFH (eg either SFHfunc, SFHburst)
          parm.names=c('mSFR','mpeak','mperiod','mskew','tau_birth','tau_screen',
                       'alpha_SF_birth','alpha_SF_screen'), # which parameters to fit for
          logged=c(T,T,T,F,T,T,F,F), # fit parameters in logged or linear space
          intervals=list(lo=c(-4,-2,-1,-0.5,-2.5,-2.5,0,0), hi=c(3,1,1,1,1.5,1,4,4)), # fitting range for parameters
          fit = 'LD', # specifies the way in which the SED should be fitted ('LD', 'optim', 'CMA', or 'check')
          mon.names=c('LP','masstot','SFRburst',paste('flux.',flux_input$filter,sep='')),
          N=length(filters), # number of observed filters
          like='norm',
          verbose=FALSE
  )
```

Now we can run an MCMC fit using LaplacesDemon (depending on machine, this might take 10-20 minutes):

```{r eval=FALSE}
set.seed(1)
LDout = LaplacesDemon(Model=ProSpectSEDlike, Data=Data,  Initial.Values=inpar,
                      control=list(abstol=0.1), Iterations=1e4, Algorithm='CHARM', Thinning=1)
```

Above we cheated by starting on the right values for the fit. In reality we would have to use an optimiser to get a reasonable starting position. **R** comes with many, but **optim** is a good simple start:

```{r, eval=FALSE}
set.seed(1)
Data$fit = 'optim'
optout = optim(par=inpar, fn=ProSpectSEDlike, Data=Data, method = 'BFGS')
```

In practice we might want to use something better like CMA to get pretty close pretty quickly (you can get this from my GitHub at asgr/cmaeshpc). Here we set the wall time to 2 minutes (which is usually enough for this many parameters).

```{r, eval=FALSE}
set.seed(1)
library(cmaeshpc)
Data$fit = 'CMA'
badpar = (Data$intervals$lo + Data$intervals$hi) / 2 #CMA is pretty tolerant of terrible initial guesses, unlike optim and LD.
CMAout = cmaeshpc(par=badpar, fn=ProSpectSEDlike, Data=Data, lower=Data$intervals$lo,
                  upper=Data$intervals$hi, control=list(trace=TRUE, maxwalltime=2))
print(CMAout$par)
```

## Exploring Results

We can check the posterior monitored distribution of stellar mass formed (masstot). This is a derived parameter given the star formation history, hence we have to monitor it as we go. Otherwise we would need to reconstruct the mass formed for all 10^4 combinations of parameters generated (which would take a long time!).

```{r, eval=FALSE}
maghist(LDout$Monitor[,"masstot"], verbose = FALSE, xlab='Stellar Mass / Msol', ylab='PDF')
abline(v=genSED$Stars$masstot, col='red')
```

We can also see what fluxes were generated by our various MCMC samples:

```{r, eval=FALSE}
magplot(flux_input$pivwave, LDout$Monitor[1,4:23], type='l', log='xy', grid=TRUE,
        xlab="Wavelength (Ang)", ylab='Flux Density / Jy')
for(i in 2:1e4){
  lines(flux_input$pivwave, LDout$Monitor[i,4:23], col=hsv(alpha=0.1))
}
points(flux_input[,c("pivwave","flux")])
magerr(flux_input$pivwave, flux_input$flux, ylo=flux_input$fluxerr)
```

## Fitting for Redshift

It is easy to add redshift as an explored parameter, given access to SED based photo-z. Basically we have to remove it from arglist and add it to parm.names, making sure we give it a reasonable range. Here we explore it in log-space:

Note since we are fitting for redshift now we can no longer pass LumDist_Mpc because this will change for every realisation.

```{r}
Data_z=list(flux=flux_input,
           arglist=list(massfunc=massfunc_snorm_trunc, agemax=agemax, Z=Zfunc_massmap_lin),
           speclib=BC03lr, 
           Dale=Dale_NormTot, 
           filtout=filtout, 
           SFH=SFHfunc, # the preferred funcitonal form of the SFH (eg either SFHfunc, SFHburst)
           parm.names=c('z', 'mSFR','mpeak','mperiod','mskew','tau_birth','tau_screen',
                        'alpha_SF_birth','alpha_SF_screen'), # which parameters to fit for
           logged=c(T,T,T,T,F,T,T,F,F), # fit parameters in logged or linear space
           intervals=list(lo=c(-3, -4,-2,-1,-0.5,-2.5,-2.5,0,0), hi=c(0, 3,1,1,1,1.5,1,4,4)), # fitting range for parameters
           fit = 'CMA', # specifies the way in which the SED should be fitted ('LD', 'optim', 'CMA', or 'check')
           mon.names=c('LP','masstot','SFRburst',paste('flux.',flux_input$filter,sep='')),
           N=length(filters), # number of observed filters
           like='norm',
           verbose=FALSE
)
```

A CMA fit returns the expected value (log10(z) = -1) pretty well:

```{r, eval=FALSE}
badpar_z = (Data_z$intervals$lo + Data_z$intervals$hi) / 2 #CMA is pretty tolerant of terrible initial guesses, unlike optim and LD.
CMAout_z = cmaeshpc(par=badpar_z, fn=ProSpectSEDlike, Data=Data_z, lower=Data_z$intervals$lo,
                    upper=Data_z$intervals$hi, control=list(trace=TRUE, maxwalltime=2))
print(CMAout_z$par)
```

## Fitting a Galaxy for Redshift

In principle if you have a real galaxy to fit and it is already in the required format of **flux_input**, you can start from this point. In this example we will do something very similar to the above, but now also fit for the redshift.

```{r}
Data=list(flux=flux_input,
          arglist=list(massfunc=massfunc_snorm_trunc, agemax=agemax, Z=Zfunc_massmap_lin), # we remove the explicit z from here
          speclib=BC03lr, 
          Dale=Dale_NormTot, 
          filtout=filtout, 
          SFH=SFHfunc, # the preferred functional form of the SFH (eg either SFHfunc, SFHburst)
          parm.names=c('z','mSFR','mpeak','mperiod','mskew','tau_birth','tau_screen', # we add z as a parameter to fit here
                       'alpha_SF_birth','alpha_SF_screen'), # which parameters to fit for
          logged=c(T,T,T,T,F,T,T,F,F), # fit parameters in logged or linear space (note z will be fitted in log space)
          intervals=list(lo=c(-3,-4,-2,-1,-0.5,-2.5,-2.5,0,0), hi=c(1,3,1,1,1,1.5,1,4,4)), # fitting range for parameters (z will be 0.001 to 10)
          fit = 'LD', # specifies the way in which the SED should be fitted ('LD', 'optim', 'CMA', or 'check')
          mon.names=c('LP','masstot','SFRburst',paste('flux.',flux_input$filter,sep='')),
          N=length(filters), # number of observed filters
          like='norm',
          verbose=FALSE
  )
```

Now we can run an MCMC fit using LaplacesDemon (depending on machine, this might take 10-20 minutes):

```{r eval=FALSE}
LDout = LaplacesDemon(Model=ProSpectSEDlike, Data=Data,  Initial.Values=c(0.1,inpar),
                      control=list(abstol=0.1), Iterations=1e4, Algorithm='CHARM', Thinning=1)
```

We can see that the redshift, fitted without any prior (or implicitly a uniform in log-space prior) is degenerate with many other parameters. In particular the total stellar mass, where higher redshifts results in more stellar mass. This is because a galaxy need to be brighter and on average more massive if placed at higher redshift to create the same amount of observed flux.

```{r, eval=FALSE}
magtri(cbind(LDout$Posterior1,SM=LDout$Monitor[5001:1e4,"masstot"]))
```

In practice we might want to use something better like CMA to get pretty close pretty quickly (you can get this from my GitHub at asgr/cmaeshpc). Here we set the wall time to 2 minutes (which is usually enough for this many parameters).

```{r, eval=FALSE}
library(cmaeshpc)
Data$fit = 'CMA'
badpar = (Data$intervals$lo + Data$intervals$hi) / 2 #CMA is pretty tolerant of terrible initial guesses, unlike optim and LD.
CMAout = cmaeshpc(par=badpar, fn=ProSpectSEDlike, Data=Data, lower=Data$intervals$lo,
                  upper=Data$intervals$hi, control=list(trace=TRUE, maxwalltime=2))
print(CMAout$par)
```

With photo-z like prior

```{r}
prior = function(parm){
return(dnorm(parm[1], mean=-1, sd=0.02, log=TRUE)) # 10^0.02 gives about 4.7% photo-z error
}
```

```{r}
Data=list(flux=flux_input,
          arglist=list(massfunc=massfunc_snorm_trunc, agemax=agemax, Z=Zfunc_massmap_lin), # we remove the explicit z from here
          speclib=BC03lr, 
          Dale=Dale_NormTot, 
          filtout=filtout, 
          SFH=SFHfunc, # the preferred functional form of the SFH (eg either SFHfunc, SFHburst)
          parm.names=c('z','mSFR','mpeak','mperiod','mskew','tau_birth','tau_screen', # we add z as a parameter to fit here
                       'alpha_SF_birth','alpha_SF_screen'), # which parameters to fit for
          logged=c(T,T,T,T,F,T,T,F,F), # fit parameters in logged or linear space (note z will be fitted in log space)
          prior=prior,
          intervals=list(lo=c(-3,-4,-2,-1,-0.5,-2.5,-2.5,0,0), hi=c(1,3,1,1,1,1.5,1,4,4)), # fitting range for parameters (z will be 0.001 to 10)
          fit = 'LD', # specifies the way in which the SED should be fitted ('LD', 'optim', 'CMA', or 'check')
          mon.names=c('LP','masstot','SFRburst',paste('flux.',flux_input$filter,sep='')),
          N=length(filters), # number of observed filters
          like='norm',
          verbose=FALSE
  )
```

```{r eval=FALSE}
LDout_prior = LaplacesDemon(Model=ProSpectSEDlike, Data=Data,  Initial.Values=c(0.1,inpar),
                      control=list(abstol=0.1), Iterations=1e4, Algorithm='CHARM', Thinning=1)
``` 
