---
title: "Spectro-Photometric fitting"
author: "Aaron Robotham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spectro-Photometric fitting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Making Mock Data

Load the libraries we will need:

```{r}
library(ProSpect)
library(LaplacesDemon)
library(foreach)
library(celestial)
library(magicaxis)
```

Load some data we want to use:

```{r}
data("BC03lr") #BC03 spectral library
data("Dale_NormTot") #Normalised Dale templates
data("pivwave") # Pivot/effective wavelengths
```

```{r}
set.seed(1)
redshift = 0.1

filters=c('FUV_GALEX', 'NUV_GALEX', 'u_SDSS', 'g_SDSS', 'r_SDSS', 'i_SDSS', 'Z_VISTA',
          'Y_VISTA', 'J_VISTA', 'H_VISTA', 'K_VISTA', 'W1_WISE' , 'W2_WISE', 'W3_WISE',
          'W4_WISE', 'P100_Herschel', 'P160_Herschel', 'S250_Herschel' , 'S350_Herschel',
          'S500_Herschel')

filtout=foreach(i = filters)%do%{approxfun(getfilt(i))}

temppiv=pivwave[pivwave$filter %in% filters,]

agemax = 13.3e9 - cosdistTravelTime(z=redshift, H0 = 67.8, OmegaM = 0.308)*1e9
```

Create our model galaxy We will be using a snorm_trunc SFH function and a massmap_lin metallicity with the following parameters:

```{r}
inpar=c(mSFR = 0, #log-space
        mpeak = 0.7, #log-space
        mperiod = 0.3, #log-space
        mskew = 0.3,
        tau_birth = 0, #log-space
        tau_screen = -0.5, #log-space
        alpha_SF_birth = 1,
        alpha_SF_screen = 3
        )
```

Simple function to view our target SFH:

```{r}
plotSFH=function(par, magemax=13.3, add=FALSE,col='black',ylim=NULL,...){
 magcurve(massfunc_snorm_trunc(age=x,mSFR=10^par[1],mpeak=10^par[2],mperiod=10^par[3],
                               mskew=par[4], magemax=magemax),0,13.8e9,add=add,col=col,
                               ylim=ylim,xlab='Age (Yr)', ylab='SFR (Msol / Yr)',...)
}
```

Let's take a look:

```{r, fig.width=8, fig.height=6}
plotSFH(inpar)
```

Now we can generate our galaxies SED:

```{r}
genSED=ProSpectSED(massfunc = massfunc_snorm_trunc,
                   mSFR = 10^inpar[1],
                   mpeak = 10^inpar[2],
                   mperiod = 10^inpar[3],
                   mskew = inpar[4],
                   tau_birth = 10^inpar[5], 
                   tau_screen = 10^inpar[6], 
                   alpha_SF_birth = inpar[7], 
                   alpha_SF_screen = inpar[8],
                   z = redshift,
                   Z = Zfunc_massmap_lin,
                   filtout = filtout,
                   Dale = Dale_NormTot,
                   speclib = BC03hr,
                   agemax = agemax,
                   magemax = agemax/1e9,
                   sparse = 1,
                   waveout = NULL
)
```

We also want a small bit of the spectrum:

```{r}
spec_input = genSED$FinalFlux[genSED$FinalFlux$wave > 3800 & genSED$FinalFlux$wave < 1e4,]
spec_input[,'fluxerr'] = spec[,'flux']*0.05
```

At this point we create our mock photometry catalogue entry with a fractional error in flux of 0.1 assumed (roughly 0.092 mag error). The photom output of **ProSpectSED** is always Jansky, which is what we want for fitting (you shouldn't fit in magnitude space!).

```{r}
phot_input = data.frame(filter=temppiv$filter, pivwave=temppiv$pivwave, flux=genSED$Photom, fluxerr=genSED$Photom*0.1)
print(phot_input)
```

## Fitting a Galaxy

In principle if you have a real galaxy to fit and it is already in the required format of **flux_input**, you can start from this point.

To speed things up we will also pre-compute the Luminosity distance to our source (since this will be the same for every iteration, and will save some computation time):

```{r}
LumDist_Mpc = cosdistLumDist(z=redshift, H0=67.8, OmegaM=0.308)
```

```{r}
Data=list(flux = phot_input,
          spec = spec_input,
          mode = 'both',
          arglist=list(z=redshift, massfunc=massfunc_snorm_trunc, agemax=agemax, magemax=agemax/1e9,
                       Z=Zfunc_massmap_lin, LumDist_Mpc=LumDist_Mpc, sparse=1, waveout=NULL),
          speclib=BC03hr, 
          Dale=Dale_NormTot, 
          filtout=filtout, 
          SFH=SFHfunc, # the preferred functional form of the SFH (eg either SFHfunc, SFHburst)
          parm.names=c('mSFR','mpeak','mperiod','mskew','tau_birth','tau_screen',
                       'alpha_SF_birth','alpha_SF_screen'), # which parameters to fit for
          logged=c(T,T,T,F,T,T,F,F), # fit parameters in logged or linear space
          intervals=list(lo=c(-4,-2,-1,-0.5,-2.5,-2.5,0.0625,0.0625), hi=c(3,1,1,1,1,1,4,4)), # fitting range for parameters
          fit = 'LD', # specifies the way in which the SED should be fitted ('LD', 'optim', 'CMA', or 'check')
          mon.names=c('LP','masstot','SFRburst',paste('flux.',flux_input$filter,sep='')),
          N=length(filters), # number of observed filters
          like='norm',
          verbose=FALSE
  )
```

```{r}
ProSpectSEDlike(inpar, Data)
```

So now we can fit both the spectra and the broadband photometry at the same time!

```{r, eval=FALSE}
library(Highlander)
badpar = (Data$intervals$lo + Data$intervals$hi) / 2
Highout = Highlander(badpar, Data=Data, likefunc=ProSpectSEDlike, Niters=c(1e3,1e3))
```
